---
title: "Combined SQL and Programming Task"
output:
  html_document:
    df_print: paged
---


I've imported and cleaned the data, but haven't continued with the task questions - these are saved here for future reference:
file:///Users/user/d6_classnotes/week_03/day_3/3_sql_and_programming_task/sql_programming_task.html

# Summary of code below:

1 Makes a link to the database

2 Links the tables employees and teams and makes a single object called employees_teams

3 Cleans the variable names of this new table

4 Creates a new object called employees_teams_clean - this stage:

a) drops  the duplicate team_id column
b) renames the column "name" as "team_name"
c) creates a new column "full_name" which concatenates first & last names





## MVP3 - source() the credentials file you created in the last step into a code chunk and use the sourced variables to create a connection to the acme_pool database

```{r}
source("~/acme_credentials.R")
```

```{r}
library(RPostgres)
library(janitor)
library(tidyverse)
library(dplyr)
```


```{r}
acme <- dbConnect(
    drv= Postgres(),
  user = Username,
  password = Password,
  dbname = Database,
  host = Host,
  port = Port,
  bigint = "numeric"
)
```

```{r}
rm(Password, Username)
```

```{r}
dbListTables(acme)
```

#MVP4 - Get the records of all employees, adding a field full_name containing the first_name and last_name of each employee separated by a space
```{r}
employees_teams <- dbGetQuery(acme,  "SELECT * FROM employees LEFT JOIN teams
ON employees.team_id = teams.id;")
```

```{r}
clean_names(employees_teams)

```

```{r}
employees_teams_clean <- 
  employees_teams %>% 
  select(-id..6) %>%
  rename("team_name"= name) %>% 
  mutate(full_name = paste(first_name, last_name, sep = " "))

employees_teams_clean
```

