---
title: "Homework week02 day2"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyverse)
```

Q1 Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.
```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
```

```{r}
dim(code_clan_tweets)
names(code_clan_tweets)
```



Q2 Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).
```{r}
code_clan_tweets %>%
  filter(is_quote == "FALSE") %>% 
  summarise(total_favourites = sum(favorite_count))  
```

Q3 Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
code_clan_tweets %>% 
  filter(is_quote == "FALSE") %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count)) %>% 
  #format(mean_retweets, nsmall = 2) %>% 
  arrange(desc(mean_retweets))
```

Q4 Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.
```{r}
code_clan_tweets %>% 
  
  # missing media_types assumed to be "text"
  mutate(media_type = if_else(is.na(media_type), "text", media_type)) %>% 
  
  group_by(media_type) %>% 
  summarise(total_likes = sum(favorite_count)) %>% 
  arrange(desc(total_likes))
```

Q5 Find the mean number of characters that a CodeClan tweet contains
```{r}
# not sure if this is the same as "display_text_width" - this seems to be zero if it's just a web link or user name? I don't use twitter, so not sure how these are handled.

#? why do some come out with NA  for code below:

#code_clan_tweets %>% 
  #mutate(number_chars = str_length(text)) %>% 
  #select(text, number_chars) %>% 
  #filter(is.na(number_chars))

# I'll just use display_text_width which has no NAs:

#code_clan_tweets %>% 
  #filter(is.na(display_text_width))

code_clan_tweets %>% 
  summarise(mean_number_chars = mean(display_text_width))

# to check - how do I round this to fewer decimal places/display it with fewer dp?

```

Q6 The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
code_clan_info <- read_csv("data/code_clan_info.csv")
```

```{r}
code_clan_tweets_and_info <-
left_join(code_clan_info, code_clan_tweets, by = "tweet_id") 

code_clan_tweets_and_info %>% 
  select(tweet_id, status_url)
        
```

Q7 From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

```{r}
# I'm not sure about how to create a tibble

codeclan_hashtags <- 
left_join(code_clan_info, code_clan_tweets, by = "tweet_id") %>%
  
filter(!is.na(hashtags)) %>% 
select(tweet_id, hashtags) %>% 
arrange(hashtags)


```

