---
title: "Homework week 10 day 1"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(broom)
library(janitor)
library(ggfortify)
```


## 1. Load the data into a dataframe `project`
```{r}
project <- read_csv("data/project_management.csv")
```


## 2. Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.
```{r}
project %>% 
  ggplot(aes(x = actual_length, y = estimated_length)) +
  geom_point()
```


## 3. Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.

Very strong correlation - between 0.8 and 0.99

```{r}
project %>%
  summarise(cor(estimated_length, actual_length))
```
## 4. Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.

```{r}
model <- lm(formula = estimated_length ~ actual_length, data = project)

model
```
## 5. Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2 value reported by the model?

Ignoring the outlier at (39,22) it seems reasonable?

r2 0.64: some of variation explained by explanitory variable?

```{r}
glance_output <- clean_names(glance(model))
glance_output
```

```{r}
tidy_output <- clean_names(tidy(model))
tidy_output
```


```{r}
project <- project %>%
  add_predictions(model) %>%
  add_residuals(model)  %>% 
  arrange(desc(actual_length))
project

project %>% 
ggplot(aes(x = actual_length)) +
  geom_point(aes(y = estimated_length)) +
  geom_line(aes(y = pred), col = "red")
```


## 6. Is the relationship statistically significant? Remember, to assess this you need to check the p-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the p-value will be reliable (don’t worry about any outlier points you see in the diagnostic plots, we’ll return to them in the extension).

p-value = 1.911284^12?

```{r}

```

