---
title: "Homework week02 day3"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyverse)
library(stringr)
```


Q1 - create list for 3 day forecast
```{r}
forecast_three_days <- list(
                        "Thursday" = list(temp = 3, chance_rain = 15,wind_speed = 6),
                        "Friday" = list(temp = 5, chance_rain = 86, wind_speed = 12),
                        "Saturday" = list(temp = 5, chance_rain = 37, wind_speed = 8)
                        )
```

Q2 - Use this list to find:

```{r}
# i The chance of rain tomorrow (Thursday)
forecast_three_days$Thursday$chance_rain
```

```{r}
# ii The wind speed on the day after tomorrow (Friday)
forecast_three_days$Friday$wind_speed
```

```{r}
# iii The temperature tomorrow (Thursday)
forecast_three_days$Thursday$temp
```



Q3 - matrix of 14 day temperature forecast
```{r}
forecast_fourteen_days <- 
matrix(c(3, 5, 5, 4, 4, 5, 5,
         6, 7, 7, 6, 6, 6, 6),
        nrow = 2, ncol = 7, byrow = TRUE)

class(forecast_fourteen_days)

forecast_fourteen_days
```



Q4 - select the following:
```{r}
# i The third day of the second week:
forecast_fourteen_days[2,3]
```

```{r}
# ii The whole second week:
forecast_fourteen_days[2,]
```

```{r}
# iii The first, third and fifth days in both weeks:
forecast_fourteen_days[,c(1, 3, 5)]
```


Q5 - Write a function that pre-appends “The answer is:” to any character input, 

while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. 

For anything else return “Invalid input”.

```{r}
# 1st attempt: test logic to be used in the function

# This seems to work, but get an error message as well as the expected response - am I doing something wrong with my if statements? Is it maybe running all of them because I've not nested them?

answer <- "A"

#prefix_answer_is <- function(answer){}

                      if(is.character(answer)) {
                        print(str_c("The answer is ", answer))
                      }
                      if(answer < 10) {
                       print(answer )
                      }
                      # I've interpreted the question as >= 10 rather than > 10

                       if(answer >= 10) {
                        print(answer/10)
                        
                        } else {print("Invalid input")
                        }
                        
                      
```


```

```{r}
# corrected answer
answer_or_divide <- function(x) {
  if (is.character(x)) {
    return(paste("The answer is: ", x))
  }

  if (is.numeric(x)) {
    if (x < 10) {
      return(x)
    } else {
      return(x / 10)
    }
  }

  return("Invalid input")
}
```


```{r}
```{r}
# 2nd attempt:
#answer <- 10

#if(is.numeric(answer){
    #if (answer < 10) {
    #print(answer)
    #}else{print(answer/10)
    #}

#Nope!
```


Q6 - function to compare two vectors
```{r}

```

