---
title: "Homework week 6 day 1"
output:
  html_document:
    df_print: paged
---
```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic_data <- read_csv("data/titanic.csv")
```

# 1) Titanic 

Calculate the conditional probabilities prob( survived ∣ crew ) and prob( survived ∣ passenger )

I'm not sure how to do this - proportion below shows that there were more passengers than crew among those who survived; not sure if this answers the question about probability.
```{r}
titanic_data %>% 
distinct(class)
```


### Add passenger/crew column
```{r}
titanic_data <- titanic_data %>% 
  mutate(passenger_crew = case_when(class %in% c("1st", "2nd", "3rd") ~ "passenger",
                                    TRUE ~ "crew")
  )
```

### Get total number of survivors
```{r}
total_number_survived <- titanic_data %>%
  filter(survived == "yes") %>% 
  nrow()
```

### Check proportion passengers/crew amongst those who survived
```{r}
titanic_data %>% 
  group_by(passenger_crew) %>% 
  filter(survived == "yes") %>% 
  summarise(number_survivors = n(),
            proportion_of_total_survivors  = round(n()/total_number_survived, 2)
            )
```



# 2) MVP

## 2.1 If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
number_possible_outcomes_one_die <- rolldie(1) %>% 
      nrow()

number_outcomes_under_five_one_die <- rolldie(1) %>% 
  filter(X1 < 5) %>% 
    nrow

# Probability of getting a number less than 5:
number_outcomes_under_five_one_die/number_possible_outcomes_one_die


  
```


## 2.2 How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
rolldie(3) %>% 
nrow()
```


## 2.3 In an experiment where we roll four dice:

### i how many outcomes are in the event total_equals_5_6_7 (i.e. the total of the four dice equals 5, 6 or 7)?
```{r}
rolldie(4) %>% 
  mutate(total = X1 + X2 + X3 +X4) %>% 
  filter(total %in% 5:7) %>% 
  nrow()

```


### ii what is the probability of this event?
```{r}
possible_outcomes_four_die <- rolldie(4) %>% 
  mutate(total = X1 + X4 + X3 +X4)

number_outcomes_four_die <- possible_outcomes_four_die %>% 
nrow()

number_outcomes_sum_5_to_7_four_die <- possible_outcomes_four_die %>% 
  filter(total %in% 5:7) %>% 
    nrow

# Probability of getting a total of 5, 6 or 7 from throwing four dice
number_outcomes_sum_5_to_7_four_die/number_outcomes_four_die
```

### iii what assumptions do we apply in answering part ii?
?




## 2.4 Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

If we assign the people the numbers 1 to 5 from tallest to shortest, we are testing for an event where they sit in the order 1-2-3-4-5 from left to right.

This is a permutation, since order matters.

Need to work out the total number of possible permutations.

### Number of permutations = 120, therefore probability = 1/120 = 0.00833333
```{r}

permutations(
  n = 5,
  r = 5
) %>% 
  as_tibble
```

## 2.5 I roll six dice:

### i what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

1/46656 = 0.00002143
```{r}
total_permutations_6_die  <- rolldie(6) %>% 
  nrow()
```


### ii what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
```{r}
six_die_all_numbers_any_order <- rolldie(6) %>% 
  filter(isin(., c(6, 5, 4, 3, 2, 1))) %>% 
  nrow()

# Probability of getting six different numbers in any order:
six_die_all_numbers_any_order/total_permutations_6_die
```


## 2.6 If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades? 

26 black cards of which 13 are spades  = 13/26 = 1/2  = 0.5

Combination; 
A: black card drawn
AND
B: spade drawn

A = 26/52 = 1/2
B = 13/26 = 1/2

P(A∩B)

eh... a bit lost now!



## 2.7 Eight patients: Alice, Bob, Carmen, Dinesh, Ella, Fergus, Gianna and Henry are waiting to see a doctor, but only five appointment slots are left today.

### i In how many different orders can the doctor see patients today?
6,720:
```{r}
patients <- c("Alice", "Bob", "Carmen", "Dinesh", "Ella", "Fergus", "Gianna",  "Henry")

permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
  as_tibble() %>% 
  rename("appt_1" = "V1",
         "appt_2" = "V2",
         "appt_3" = "V3",
         "appt_4" = "V4",
         "appt_5" = "V5")
```

### ii What is the probability of Gianna and Danesh both being seen today?

```{r}
all_permutations_appointments  <- 
  permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
  as_tibble()

count_permutations_appointments  <- 
  permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
nrow()

# g_and_d_permutations_appointments  <- 
# all_permutations_appointments %>% 
# filter(isin(c("Gianna","Danesh")))

```





