---
title: "Homework week 5 day 2"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(CodeClanData)
library(ggthemes)

```

```{r}
# Create colour scheme
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```

# MVP
## 4.1 Part 1
### 1. Take the data in the dataset **qb_revenue_breakdown** and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  scale_fill_manual(values = col_scheme) +
  labs(
    x = "\n Year",
    y = "Revenue \n",
    title = "Revenue by year",
    subtitle = "Split by product type \n",
    fill = "" 
  )
```

### 2. Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
qb_monthly_sales <-
  qb_monthly_sales
```

```{r}
qb_monthly_sales_flagship <-
  qb_monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") 
  

```

```{r}
ggplot(qb_monthly_sales_flagship) +
  aes(x = Date, y= Amount) +
  geom_line() +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  scale_colour_manual(values = col_scheme) +
  labs(
    x = "\n Date",
    y = "Value \n",
    title = "Kwikbit Unit 1 Sales",
    subtitle = "For the 12 months to Jul 2019 \n"
  )
  
```

### 3. Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
qb_monthly_sales_costs_rev <-
  qb_monthly_sales %>% 
  filter(Cashflow %in% c("Total Revenue", "Cost") )
```

```{r}
ggplot(qb_monthly_sales_costs_rev) +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  scale_y_continuous(breaks = c(1000, 1500, 2000, 2500, 3000),  labels = scales::comma) +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  scale_colour_manual(values = col_scheme) +
  labs(
    x = "\n Date",
    y = "Value \n",
    title = "Cost and total revenue",
    subtitle = "For the 12 months to Jul 2019 \n",
    colour = ""
  )
```

### 4.Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in **qb_competitors**.

```{r}
qb_competitors <- 
  qb_competitors 
```

```{r}
ggplot(qb_competitors) +
  aes(x = Year, y =  Revenue, colour = Company) +
  scale_y_continuous(labels = scales::comma) +
  geom_area() +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  scale_colour_manual(values = col_scheme) +
   labs(
    x = "\n Year",
    y = "Revenue \n",
    title = "Annual sales of personal fitness trackers",
    subtitle = "For competitors 1 - 4 \n",
    colour = ""
   )
```

### 5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(qb_competitors) +
  aes(x = Year, y =  Revenue, colour = Company) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 8)) +
  scale_colour_manual(values = col_scheme) +
  labs(
    x = "\n Year",
    y = "Revenue \n",
    title = "Annual sales of personal fitness trackers",
    subtitle = "For competitors 1 - 4 \n",
    colour = ""
   )
```


### 6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in **qb_device_data**.

* Not sure how to plot the time component of this

* ?what is the data showing - is this from a person wearing all the devices for 24 hours?

* Am adding a numeric  "time", showing minutes as a decimal of hours, to have a look, tho I think this is unnecessary and over-complicated!

* I've used a free y scale here just to keep things visible;  this is misleading - although the graphs for different products have similar shapes, when you look at the scale, the number of steps recorded has some large variances
```{r}
qb_device_data_with_time <- qb_device_data %>%
  mutate(mins_numeric = mins/60) %>% 
  mutate(time = hours + mins_numeric)
  
```

```{r}
qb_device_data_with_time %>% 
  arrange(desc(counts)) %>% 
  head(100)
```



```{r}
ggplot(qb_device_data_with_time) +
  aes(x = time, y  = counts, colour = id, ) +
  geom_col(show.legend = FALSE) +
  #ylim(0,3000) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal(base_size = 8) +
  theme(axis.title = element_text(size = 6)) +
  theme(legend.text = element_text(size = 6)) +
  scale_colour_manual(values = col_scheme ) +
  facet_wrap(device ~ id, scales = "free_y") +
  labs(
    x = "\n Time (hours)",
    y = "Steps counted \n",
    title = "Steps counted over time",
    subtitle = "For 5 people using quikbit and the products of competitors 1 - 3 \n"
   )
```

