---
title: "Homework week02 day1"
output: html_notebook
---
```{r}
library(tidyverse)
library(dplyr)
library(janitor)
```


```{r}
card_base <- read_csv("data/CardBase.csv")
customer_base <- read_csv("data/CustomerBase.csv")
fraud_base <- read_csv("data/FraudBase.csv")
transaction_base <- read_csv("data/TransactionBase.csv")
```

# 1.1 Joins
## Q1 - clean variable names
```{r}
card_base <- clean_names(card_base)
customer_base <- clean_names(customer_base)
fraud_base <- clean_names(fraud_base)
transaction_base <- clean_names(transaction_base)
```

## Q2 - join card details to customer details
This returns the same number of rows as the card_base table - this is what we expect, as there should be a customer attached to every credit card
```{r}
left_join(card_base, customer_base, by = "cust_id") 
```

## Q3 - join fraud details to transaction details
This returns the same number of rows as the transactions table, because fraudulent transactions are a subset of all transactions;

although this type of join returns all records from both tables, it is not possible to have a fraud transaction which doesn't also appear in a table of all transactions. Any extra rows would suggest an error in the fraud table.
```{r}
full_join(fraud_base, transaction_base, by = "transaction_id") 
```

## Q4 - join card details to transaction details, returning only one instance of the card details
```{r}
semi_join(card_base, transaction_base, by = c("card_number"= "credit_card_id"))
```


# 1.2 tidyr
```{r}
library(tidyr)
```

```{r}
hat_observations <- read_csv("data/hat_observations.csv")
```

## Q5 -separate one column into two
```{r}
hat_observations_new_cols <- hat_observations %>% 
  separate(
    col = observation,
    into = c("hat_colour", "hat_type"),
    sep = "\\,"
  )
```

## Q6 - use unite to create one column from three
Most berets were observed on 18/06/2018
```{r}

hat_observations_new_cols %>% 
  unite(
    col = "date",
    c(day, month, year),
    sep = "/"
  ) %>% 
    
  mutate(date = as.Date(date, "%d/%m/%Y")) %>% 
    
  filter(hat_type == "beret") %>% 
  group_by(date, hat_type) %>% 
  summarise(total_observations = sum(observation_count)) %>% 
  arrange(desc(total_observations)) 
  #slice_max(total_observations)
    
    
```

# 2.1 Joins
join 4 datasets
```{r}
complete_finance_data <- 
  left_join(customer_base, card_base, by = "cust_id" ) %>% 
  left_join(transaction_base, by = c("card_number" = "credit_card_id")) %>%
  left_join(fraud_base, by = "transaction_id") 
  
```

```{r}
  complete_finance_data %>% 
  filter(fraud_flag == 1) %>% 
  select(
    card_number,
    card_family,
    credit_limit,
    cust_id,
    age,
    customer_segment,
    customer_vintage_group,
    transaction_id,
    transaction_date,
    transaction_value,
    transaction_segment,
    fraud_flag
  ) %>% 
  arrange(desc(transaction_value))
  
```

