---
title: "NYC Citi Bike Report"
output:
  html_document:
    df_print: paged
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
library(lubridate)
library(tibble)
library(tidyverse)
library(tsibble)
library(tsibbledata)
library(slider)

```


```{r}
nyc_bikes_df <- nyc_bikes %>% 
  mutate(
    start_date = as_date(start_time),
    stop_date = as_date(stop_time),
    start_year = year(start_time),
    start_quarter = quarter(start_time),
    start_month = month(start_time,
                        label = TRUE,
                        abbr = FALSE),
    start_weekday = wday(start_time,
                    week_start = 1,
                    label = TRUE,
                    abbr = FALSE),
    trip_duration = (stop_time - start_time),
    age = start_year - birth_year,
    #hire_duration_moving_average = slide_period(
    #  hire_duration,
    #  ~ mean(., na.rm = TRUE),
    #  .before = 100,
    #  .after = 100,
    #.complete = TRUE
    #)
  )
```

# Overview

This report uses a sample of NYC Citi Bike usage data to assess the following questions:

  * What is the pattern of bike hires over time (e.g. within a year, month, week, or day)?
  
  * Do bike hire patterns differ between bike rider demographics? (e.g. gender, type of trip, age)
  
  * Any other insights?
  
  
# Background

## About NYC Citi Bike

Citi Bike is New York City’s bike share system. Launched in 2013, it has over 140,000 members, and in July 2020 it recorded its 100 millionth trip.

**Website:**
https://www.citibikenyc.com/about



## About the dataset

This report uses a sample which details the usage of 10 bikes throughout 2018.

The dataset is in the public domain, and contains no person-identifiable information.

**Website:**
https://rdrr.io/github/tidyverts/tsibbledata/man/nyc_bikes.html



# Analysis of the data

## Number of trips per day

Add commentary: trend peaks in summer, highest single trips date was ??, 

```{r}
nyc_bikes_df %>%
  index_by(start_date) %>%
  summarise(number_of_trips = n()) %>% 
ggplot() +
  aes(x = start_date, y = number_of_trips) +
  geom_line(colour = "steel blue") +
  #theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Number of trips per day",
    x = "",
    y = "Number of trips \n"
  )
```


## Number of trips - gender of user
Add commentary: where gender was recorded, there were many more trips by males:

+--------+-----------------+
| Gender | Number of trips |
+--------+-----------------+
| Female | 930             |
+--------+-----------------+
| Male   | 3,069           |
+--------+-----------------+

```{r}
#nyc_summ_gender <- tibble( nyc_bikes_df %>%
#  filter(gender!= "Unknown") %>% 
#  group_by(gender) %>% 
#  summarise(number_of_hires = n()) )
#nyc_summ_gender <- nyc_summ_gender %>% 
#  group_by(gender) %>% 
#  summarise(number_of_hires = sum(number_of_hires))
  
nyc_bikes_df %>%
  filter(gender!= "Unknown") %>% 
  index_by(start_month) %>%
  group_by(gender) %>% 
  summarise(number_of_trips = n()) %>% 
ggplot() +
  aes(x = start_month, y = number_of_trips) +
  geom_col(fill = "steel blue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~gender) +
  labs(
   # title = "Number of trips per month",
    x = "\n 2018",
    y = "Number of trips \n"
  )
```
## ?check age - is there one 49 year-old who keeps hiring or just general MAMIL behaviour
```{r}


nyc_bikes_df %>%
  filter(age < 130) %>% 
  index_by(age) %>%
  group_by(gender) %>% 
  summarise(number_of_hires = n()) %>% 
ggplot() +
  aes(x = age, y = number_of_hires) +
  geom_col(fill = "steel blue") +
  theme_minimal() 


```



# Conclusions and key insights

?address gender imbalance



