---
title: "NYC Citi Bike Report"
output:
  html_document:
    code_folding: hide
---

```{r, warning = F, message = F}
library(lubridate)
library(tibble)
library(tidyverse)
library(tsibble)
library(tsibbledata)
library(slider)
library(leaflet)

```


```{r, warning = F, message = F}
# Add derived columns to dataset

nyc_bikes_df <- nyc_bikes %>% 
  mutate(
    start_date = as_date(start_time),
    stop_date = as_date(stop_time),
    start_year = year(start_time),
    start_quarter = quarter(start_time),
    start_month = month(start_time,
                        label = TRUE,
                        abbr = FALSE),
    start_weekday = wday(start_time,
                    week_start = 1,
                    label = TRUE,
                    abbr = FALSE),
    start_hour = hour(start_time),
    start_minute = minute(start_time),
    start_hour_numeric = (start_hour + (start_minute/60)),
    trip_duration = (stop_time - start_time),
    age = start_year - birth_year,
    age_group = case_when(age  >= 65 ~ "65+",
                          age  >= 55 ~ "55 - 64",
                          age  >= 45 ~ "45 - 54",
                          age  >= 35 ~ "35 - 44",
                          age  >= 25 ~ "25 - 34",
                          age  >= 18 ~ "18 - 24",
                          TRUE ~ NA_character_
    )
  )
```

# Overview

This report uses a sample of NYC Citi Bike usage data to assess the following questions:

  * What is the pattern of bike hires over time?
  
  * Do bike hire patterns differ between bike rider demographics?
  
  * What is the geographical spread of the start points of bike hires?
  
  * Any other insights?
  
  
# Background

## About NYC Citi Bike

Citi Bike is New York City’s bike share system. Launched in 2013, it has over 140,000 members, and in July 2020 it recorded its 100 millionth trip.

**Website:**
https://www.citibikenyc.com/about



## About the dataset

This report uses a sample which details the usage of 10 bikes throughout 2018.

The dataset is in the public domain, and contains no person-identifiable information.

**Link:**
https://rdrr.io/github/tidyverts/tsibbledata/man/nyc_bikes.html



# Analysis of the data

## Pattern of bike hires over time

**Number of trips per day in 2018**

As might be expected, there are more trips per day in summer than in winter.

This peaked at 47 trips on the 31st of July, while there were a number of dates in Jan-Apr and Nov-Dec when only one trip was recorded.

The moving average shown in red suggests that usage plateaus through the summer months - it might be worth investigating further to see if this has any relation to the availability of bikes (for example, if demand for trips exceeds the number of bikes available).

```{r, warning = F, message = F}
 #Date with HIGHEST number of trips
  #nyc_bikes_df %>%
  #index_by(start_date) %>%
  #summarise(number_of_trips = n()) %>% 
  #arrange(desc(number_of_trips)) %>% 
  #head(1)

# Dates with LOWEST number of trips
  #nyc_bikes_df %>%
  #index_by(start_date) %>%
  #summarise(number_of_trips = n()) %>% 
  #filter(number_of_trips == 1) 
```


```{r, warning = F, message = F}

# Source data for trips per day
nyc_bikes_df %>%
  index_by(start_date) %>%
  summarise(number_of_trips = n()) %>% 
  mutate(
    trips_moving_average = slide_dbl(
      number_of_trips,
      ~ mean(., na.rm = TRUE),
      .before = 75,
      .after = 75,
      .complete = FALSE
    )
  ) %>% 
ggplot() +
  aes(x = start_date, y = number_of_trips) +
  geom_line(colour = "light grey") +
  geom_line(aes(y = trips_moving_average), colour = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Number of trips per day in 2018 \n",
    x = "",
    y = "Number of trips \n"
  )
```

**Number of trips by time of day**

Again, this follows a pattern you might expect - there are peaks around typical commuting hours, with usage rising sharply before the morning peak, and tailing off more gradually in the evening.


```{r}
# source data for trips by time of day
nyc_bikes_df %>%
  index_by(start_hour_numeric) %>%
  summarise(number_of_trips = n()) %>% 
  mutate(
    trips_moving_average = slide_dbl(
      number_of_trips,
      ~ mean(., na.rm = TRUE),
      .before = 75,
      .after = 75,
      .complete = FALSE
    )
  ) %>% 
ggplot() +
  aes(x = start_hour_numeric, y = number_of_trips) +
  geom_line(colour = "light grey") +
  geom_line(aes(y = trips_moving_average), colour = "red") +
  scale_x_continuous(breaks = 1:24) +
  theme_minimal() +
  labs(
    title = "Number of trips by time of day \n",
    x = "\n Time (24h)",
    y = "Number of trips \n"
  )
```

## Demographic analysis 

**Total number of trips in 2018 by gender**

The first figure to note is that more than three times as many trips were made by males than females:





*Total number of trips in 2018 by gender:*

+--------+-----------------+
| Gender | Number of trips |
+--------+-----------------+
| Female | 930             |
+--------+-----------------+
| Male   | 3,069           |
+--------+-----------------+

Note that records where gender was recorded as "Unknown" have been excluded, as I'm not familiar enough with how the data was recorded to be sure if these represent a deliberate choice by the user, or if they represent missing data. 

```{r, warning = F, message = F}
nyc_summ_gender <- tibble( nyc_bikes_df %>%
  filter(gender!= "Unknown") %>% 
  group_by(gender) %>% 
  summarise(number_of_hires = n()) )

nyc_summ_gender <- nyc_summ_gender %>% 
  group_by(gender) %>% 
  summarise(number_of_hires = sum(number_of_hires))
```


**Pattern of bike hires over time by gender**

If we repeat the analysis on usage over the year and by time of day but show separate graphs by gender, we see a broadly similar result - while trips at the start of the year appear higher for females, it should be noted that the number of trips per day are in the low single figures.

This would merit further statistical analysis using a larger sample with more female respondents.

```{r}
# Source data for trips per day
nyc_bikes_df %>%
  filter(gender!= "Unknown") %>% 
  index_by(start_date) %>%
  group_by(gender) %>% 
  summarise(number_of_trips = n()) %>% 
  mutate(
    trips_moving_average = slide_dbl(
      number_of_trips,
      ~ mean(., na.rm = TRUE),
      .before = 75,
      .after = 75,
      .complete = FALSE
    )
  ) %>% 
ggplot() +
  aes(x = start_date, y = number_of_trips) +
  geom_line(colour = "light grey") +
  geom_line(aes(y = trips_moving_average), colour = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Number of trips per day in 2018 \n",
    x = "",
    y = "Number of trips \n"
  ) +
  facet_grid(~gender)
```


```{r}
# source data for trips by time of day
nyc_bikes_df %>%
  filter(gender!= "Unknown") %>% 
  index_by(start_hour_numeric) %>%
  group_by(gender) %>% 
  summarise(number_of_trips = n()) %>% 
  mutate(
    trips_moving_average = slide_dbl(
      number_of_trips,
      ~ mean(., na.rm = TRUE),
      .before = 75,
      .after = 75,
      .complete = FALSE
    )
  ) %>% 
ggplot() +
  aes(x = start_hour_numeric, y = number_of_trips) +
  geom_line(colour = "light grey") +
  geom_line(aes(y = trips_moving_average), colour = "red") +
  scale_x_continuous(breaks = c(3, 6, 9, 12, 15, 18, 21, 24)) +
  theme_minimal() +
  labs(
    title = "Number of trips by time of day \n",
    x = "\n Time (24h)",
    y = "Number of trips \n"
  ) +
  facet_grid(~gender)
```

**Number of trips by age of user**

The people who made trips in this sample range in age from 19 to 71 (excluding a couple of people aged over 130, I've assumed these are data entry errors!)

The majority of trips were made by people in the 25-34 age bracket.



```{r, warning = F, message = F}
nyc_summ_age_group <- tibble( nyc_bikes_df %>%
  filter(age < 130) %>% 
  group_by(age, age_group) %>% 
  summarise(number_of_trips = n()) )

nyc_summ_age <- nyc_summ_age_group %>% 
  group_by(age) %>% 
  summarise(number_of_trips = sum(number_of_trips)) 

#nyc_summ_age %>% 
#  slice_max(age) 

#nyc_summ_age %>% 
#  slice_min(age) 


```

```{r}
ggplot(nyc_summ_age_group) +
  aes(x = age_group, y = number_of_trips) +
  geom_col(fill = "steel blue") +
  theme_minimal() +
  labs(
    title = "Number of trips in 2018",
    subtitle = "By age group \n",
    x = "\n Age group",
    y = "Number of trips \n"
  )
```


## Geographical spread of the start points of bike hires

```{r}

```


# Conclusions and key insights

?address gender imbalance
