---
title: "Cleaning & Dplyer Homework"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
movie_data<- read_csv("data/movies.csv")
```

#Q1: Initial investigationof data and ideas for further investigation & cleaning:
 Cleaning:
 - original_title and title contain special chars, e.g. $, #
 - some taglines contain full stop
 - The varable name "title" is also the name of an R function
 - "Overview" column could be a nightmare - a long string of free text
 
 Further investigations:
 - Add columns to compare budget and revenue
 
```{r}
head(movie_data, 10)
names(movie_data)


```

```{r}
movie_data %>% 
  distinct(original_title, title, .keep_all = TRUE)
```

```{r}
movie_data %>% 
  group_by(original_title, title) %>% 
  summarise(sum(budget))
```

#Q2: create subset of 3 variables:
```{r}
movies_selected <- select(movie_data, title, runtime, budget)
```

#Q3: Count number of missing values in each column:
title has o rows; runtime 2; budget 0
```{r}
# Homework review: could group these as one qry
movies_selected %>% 
summarise(count = sum(is.na(title)))
```
Check visually:
```{r}
movies_selected  %>% 
  filter(is.na(title))
```

#Q4: convert 0 runtime to NA:
```{r}
movies_runtime <- 
movies_selected %>% 
  mutate(runtime = na_if(runtime, 0))
```
Check visually: 
```{r}
#check for old value 
movies_runtime %>% 
  filter(runtime == 0)

#check for new value 
movies_runtime %>% 
  filter(is.na(runtime))
```

Count missing runtime values:
There are 37
```{r}
movies_runtime %>% 
  summarise(count = sum(is.na(runtime)))
```

#Q5: impute missing values with mean:
```{r}
movies_imputed <- 
  movies_runtime %>% 
  mutate(runtime = coalesce(
                    runtime,
                    mean(runtime, na.rm = TRUE)
  ))
  
```


check for missing run times:
```{r}
movies_imputed %>% 
  filter(runtime == 0 | is.na(runtime))
```


#Q6: slice_min & slice_max to show 10 shortest & longest runtimes:
10 longest:
```{r}
movies_imputed %>% 
  slice_max(runtime, n = 10)
```

10 shortest:
```{r}
movies_imputed %>% 
  slice_min(runtime, n = 10)
```

#Q7: overwrite any budget less than $100 with median budget:
```{r}
movies_imputed %>% 
  mutate(
    budget = if_else(
              budget < 100,
              median(budget),
              budget
    )
  )
```

