---
title: "Homework week 10 day 3"
output: html_notebook
---

```{r}
library(car)
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
library(devtools)
```

```{r}
kc_house_data <- read_csv("data/kc_house_data.csv")
```

# Clean data

## Check contents of `waterfront`
```{r}
wf <- kc_house_data %>% 
  distinct(waterfront)

wf
```

## Convert `waterfront` to true/false, drop na, drop unwanted or redundant fields
```{r}
kc_house_data_trim <- kc_house_data %>% 
  drop_na() %>%
  mutate(waterfront = ifelse(waterfront == 1, TRUE, FALSE)) %>% 
  # these steps added after homework review
  mutate(condition = as_factor(condition)) %>%
  mutate(grade = as_factor(grade)) %>% 
  #****************************************
  select(-c(id, date, zipcode, sqft_living15, sqft_lot15))
```

## Check content of `yr_renovated` -  do I want to make this logical?

I'm not going to use this field to make a logical variable - I'm not sure it's complete for all records, or how "renovated" has been defined.

e.g. 

* only 914 of 21,613 houses have ever been renovated according to the data

* of almost 5,000 properties built before 1950, only 12% had ever been renovated by 2014/15 
These numbers don't feel right.


I should probably model it anyway though to see what it does... but dropping it for the moment

**Number renovated/not renovated:**
```{r}
kc_house_data_trim %>% 
  mutate(renovated = ifelse(yr_renovated == 0, FALSE, TRUE)) %>% 
  group_by(renovated) %>% 
  summarise(count = n()) 

```


**Number built before 1950 and never renovated:**
```{r}
kc_house_data_trim %>% 
  filter(yr_built < 1950) %>% 
  mutate(renovated = ifelse(yr_renovated == 0, "Never renovated", "Renovated")) %>% 
  group_by(renovated) %>% 
  summarise(count = n()) 
```


**Number renovated/not renovated by build year:**
```{r}
kc_house_data_trim %>% 
  mutate(renovated = ifelse(yr_renovated == 0, "Never renovated", "Renovated")) %>% 
  group_by(yr_built, renovated) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = renovated, values_from = count) %>% 
  arrange(yr_built)
```


**Drop yr_renovated**
```{r}
kc_house_data_trim <- kc_house_data_trim %>%
  select(-yr_renovated)
  
```


## condition and grade

These are categorical ordinal data types - the numbers have a hierarchy which means something regarding the quality of the house



# Aliased variables

This tells us that **sqft_living = sqft_above + sqft_basement**

Not totally sure which to remove:

* Do I just keep `sqft_living` because that's the total?

* Or keep the other 2 to see if e.g. amount of basement space is important, whilst the total `sqft_living` can be derived?


I'll drop `sqft_living` for the moment

**Check for alias**
```{r}
alias(price ~., data = kc_house_data_trim)
```


**Visual check of `sqft` fields in data**
```{r}
kc_house_data_trim %>% 
  mutate(check_sqft_living_minus_above = sqft_living - sqft_above) %>% 
  select(sqft_living, sqft_above, sqft_basement, check_sqft_living_minus_above) %>% 
  head(10)
```
**Drop sqft_living**
See model building steps below - I've put this back in
```{r}
#kc_house_data_trim <- kc_house_data_trim %>%
#  select(-sqft_living)
  
```


# Look for first predictor

```{r}
#kc_house_data_trim %>% 
#  ggpairs(aes(colour = waterfront, alpha = 0.5))
```

Not sure how to do that bit - just going to start with relationship between `price` and `sqft_above`

## solution from homework review
```{r}
houses_tidy_numeric <- kc_house_data_trim %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- kc_house_data_trim %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- kc_house_data_trim$price

ggpairs(houses_tidy_numeric, proportions = 10)
```


## model1 - above ground sq ft
```{r}
model1 <- lm(price ~ sqft_above, data = kc_house_data_trim)

model1
```
So price = 59953.2 + 268.5*sqft_above

```{r}
summary(model1)
```
## model1 interpretation:
Above ground sqft is statistically significant, and explains 36% of varaiation in price


## model2 - above ground sq ft and basement sq ft

```{r}
model2 <- lm(price ~ sqft_above + sqft_basement, data = kc_house_data_trim)

model2
```

```{r}
summary(model2)
```

## model2 interpretation:
This is still statistically significant, and now explains 49% of variance.

Maybe I should've just kept total sqft, at least to compare to this

I'm going to put it back in to try this


## model3 - sqft_living

```{r}
model3 <- lm(price ~ sqft_living, data = kc_house_data_trim)

model3
```
```{r}
summary(model3)
```
## model3 interpretation:
This has very similar results to using above ground & basement seperately, so for simplicity I'll continue with model3 and drop the other two fields as redundant.

```{r}
kc_house_data_trim <- kc_house_data_trim %>%
  select(-c(sqft_above, sqft_basement))
```



## model4 - sqft &  condition
```{r}
model4 <- lm(price ~ sqft_living + condition, data = kc_house_data_trim)

model4
```
```{r}
summary(model4)
```
## model4 interpretation: 
This doesn't seem to add anything to my model?  Statistically significant, but still 49% of variation explained


## model5 - sqft &  waterfront

```{r}
model5 <- lm(price ~ sqft_living + waterfront, data = kc_house_data_trim)

model5
```
```{r}
summary(model5)
```

```{r}
par(mfrow = c(2, 2))

plot(model5)
```


## model5 interpretation:
Statistically signficant, 53% of variation explained

### Next predictor:
Look at residuals: Not sure which to look at, or how to do it with numeric fields
```{r}
kc_house_data_remaining_resid <- kc_house_data_trim %>% 
  add_residuals(model5) %>% 
  select(-c("price", "sqft_living", "waterfront"))

#kc_house_data_remaining_resid %>% 
#  ggpairs(aes(colour = bedrooms))
```


## ? try PCA
```{r}
kc_house_data_numeric <- kc_house_data_trim %>%
  select(-waterfront)
```

```{r}
kc_house_pca <- prcomp(kc_house_data_numeric, center = TRUE, scale. = TRUE)
```

```{r}
summary(kc_house_pca)
```

```{r}
install_github("vqv/ggbiplot")

```

```{r}
# ggbiplot(kc_house_pca, obs.scale = 1, var.scale = 1)
```

## model6 - add condition

```{r}
model6 <- lm(price ~ sqft_living + waterfront + condition, data = kc_house_data_trim)

model6
```

```{r}
summary(model6)
```

