---
title: "Homework week 6 day 3"
output:
  html_document:
    df_print: paged
---

## 1. Load house sales data and explore it

```{r}
library(tidyverse)
library(janitor)
library(infer)
library(fastGraph)
```

```{r}
ames_data <- read_csv("data/ames.csv") %>% 
  clean_names()
```

```{r}
head(ames_data)
names(ames_data)
```

```{r}
ames_data %>% 
  ggplot() +
  aes(x = year_built) +
  scale_y_continuous(labels = scales::comma) +
  geom_histogram(col  = "white", bins = 50) 

ames_data %>% 
  ggplot() +
  aes(x = year_built, y = sale_price) +
  scale_y_continuous(labels = scales::comma) +
  geom_col() 
  
```


```{r}
ames_data %>% 
  filter(year_built >= 2000) %>% 
  group_by(year_built, bldg_type) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = bldg_type, values_from = count)
```


## 2. Plot a histogram of sale_price. Is the distribution normal?

It looks like a reasonable maths for a normal distribution, but there are a lot of values above the line at the lower end of the price scale.

```{r}
ames_stats <- ames_data %>% 
  summarise(
    num = n(),
    mean = mean(sale_price),
    sd = sd(sale_price)
  )

ames_data %>% 
  ggplot() +
  aes(x = sale_price) +
  geom_histogram(aes(y = ..density..), col = "white", fill = "Steel Blue", bins = 50)  +
  #facet_wrap(~condition_1)  +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    x = "\n Sale Price",
    y = "Density \n"
  ) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = ames_stats$mean,
      sd = ames_stats$sd
    ),
    col = "red"
  )
  
```

## 3. Calculate the population parameter mean of sale_price. Store it in a variable, as we’ll need this value later when we check if each CI contains the parameter value or not.

```{r}
pop_parameter <- ames_data %>% 
  summarise(mean_sale_price = mean(sale_price),
  ) %>% 
  pull()
```

## 4. Now we are going to:

* take 1000 random samples each of size 100 from sale_price in ames

* create a boostrap sampling distribution of mean(sale_price) from each 100-observation sample (use 1000 reps)

* calculate the 95% CI for each bootstrap distribution

* store the upper and lower bounds of the CI in a tibble called results

Taking code from notes:
```{r}
results <- tibble()
for (sample_num in 1:1000){

  # get a new 100-observation random sample from ames
  this_sample <- ames_data %>%
    rep_sample_n(size = 100, reps = 1) %>%
    ungroup() %>%
    select(sale_price)
  
  # create a bootstrap distribution of mean(sale_price)
  # use 1000 reps
  this_bootstrap <- this_sample %>%
    specify(response = sale_price) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "mean")

  # get 95% CI for this bootstrap distribution
  this_ci <- this_bootstrap %>%
    get_ci(level = 0.95, type = "percentile")
  
  # prepare new row for results, storing CI bounds
  this_result <- c(
    sample_num = sample_num, 
    lower = this_ci %>% 
      pull(lower_ci), 
    upper = this_ci %>%
      pull(upper_ci)
  )
  
  # add this_result to growing tibble of results
  results <- results %>%
    bind_rows(this_result)
}                            
```


## 5. Pipe results to mutate() to add a new variable popn_param_in_ci which will be TRUE if the population parameter we calculated earlier is between lower and upper, and FALSE otherwise.

```{r}
results <- results %>% 
  mutate(popn_param_in_ci = if_else(
                                    (pop_parameter >= lower &
                                       pop_parameter <= upper),
                                    TRUE,
                                    FALSE)
  )
```


## 6.Perform an appropriate summarise() to determine the proportion of CIs that contain the population parameter.

Our confidence level was 95% (equivalent to a proportion of 0.95). Is your proportion close to that value?

```{r}
 100 * nrow(results %>% 
        filter(popn_param_in_ci == TRUE))/nrow(results)
```


## 7. Have a think about what all this means, make a list of any points you don’t understand.

I'd need to go through section 4 again - didn't really follow what the code was doing.


