---
title: "week01_day2_dplyer_homework"
output: html_notebook
---
call tidyverse
```{r}
library(tidyverse)

```

Read in data to starbucks_data
```{r}
starbucks_data <- read_csv("data/starbucks_drinkMenu_expanded.csv")
```


Q1: Basic data checks:
Dim: 242 obvs of 18 variables
Names: 2 text variables relating to name and preparation method, remaining are numeric (measurement or %) relating to nutritional content
```{r}
dim(starbucks_data)
names(starbucks_data)
head(starbucks_data, 10)

#corrected version:
starbucks_data %>% 
select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  filter(is.na(Calories))


```

Q2:select variables, drop NA observations:
```{r}
calory_data <- select(starbucks_data, Beverage_category, Beverage, Beverage_prep, Calories )
# need to check how to ignore na
calory_data_no_nulls <- filter(calory_data, Calories != "NA")

head(calory_data_no_nulls, 100)
head(calory_data, 100)
dim(calory_data)
dim(calory_data_no_nulls)

```

Q3: add column with variance to 135 cals:
```{r}
 data_var_to_135_cals <- mutate(calory_data_no_nulls, calorie_diff = Calories - 135)

# corrected version: Note pipe means you don't have to repeat data source name
starbucks_data %>% 
  mutate(calorie_diff = Calories- 135)

```


Q4: Summarise mean calories by bevarage type:

Three beverages with highest cals: Smoothies, Frappuccino® Blended Coffee, Signature Espresso Drinks
Three lowest: Coffee, Shaken Iced Beverages, Classic Espresso Drinks
```{r}
cals_by_bev_type <- select(data_var_to_135_cals, Beverage_category, Calories) %>% 
  group_by(Beverage_category)

summarise(cals_by_bev_type, mean(Calories)) 
# Check how to arrange results by mean(Calories)
 
# corrected version - note some could be saved as code and called for head/tail versions

starbucks_data %>% 
  group_by(Beverage_category)  %>% 
  summarise(avg_cals = mean(Calories)) %>% 
  arrange (avg_cals) %>% 
  head(3)


```

Q5: Average by prep type:
Note that Beverage_prep category would benefit from refining - e.g. categories like "Tall" and "Tall Nonfat Mailk" suggest perhaps splitting into two variables, size and milk/soymilk type
```{r}
cals_by_bev_prep <- select(data_var_to_135_cals, Beverage_prep, Calories) %>% 
  group_by(Beverage_prep)

summarise(cals_by_bev_prep, mean(Calories)) 
# Check how to arrange results by mean(Calories)
```

Q6: Beverage_prep calories vs all-beverage average
2% Milk, Grande Nonfat Milk and Soymilk have >average calorie content
```{r}
# get all-beverage average
average_cals <- starbucks_data %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
pull()

summarise(cals_by_bev_prep, mean(Calories)) # %>% 
  # check how to filter for cals > average_cals
  # filter(cals_by_bev_prep, mean(Calories) >= average_cals)



```

Q7: Plain ol' coffee, averaging at 4.25 calories, is the best option for those counting calories
